\documentclass[12pt]{article}
%\usepackage{palatino, amsfonts, amsmath, fancyheadings}
\usepackage{amsfonts, amsmath, fancyhdr,lastpage,multirow}
\usepackage{amsthm}
\usepackage{graphicx}
\pagestyle{fancy}

\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

%% For marking tables

\usepackage{tikz}
\usetikzlibrary{fit,shapes.misc}

\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (0,1.5ex) {};%
}
\newcommand\markbottomright[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0,0) {};%
    \tikz[overlay,remember picture,thick,dashed,inner sep=3pt]
        \node[draw, rectangle,fit=(marker-#1-a.center) (marker-#1-b.center)] {};%
}
%% End marking tables


\topmargin = -0.8 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\textheight = 9 in
\textwidth = 6.75 in

\newtheorem{theorem}{Theorem}   
\newtheorem{result}{Result}
\newtheorem{definition}{Defintion}  
\newcounter{problem}
\newcounter{solution}
\def\bibite{$\div \hskip-0.77em \times$}
\def\endex{\bibite \kern-3pt \bibite \kern-3pt \bibite}
\def\theproblem{\arabic{problem}}
\newenvironment{problem}
     {\refstepcounter{problem}%
    \medskip\noindent{\bfseries Problem} {\bfseries \theproblem}
      }
%   {\vskip-1.5em  \hphantom{a} \hfill \endex
%     \smallskip}

\newenvironment{solution}
     { 
         % \refstepcounter{solution}%
         \setlength{\parskip}{0.5em}
         \setlength{\parindent}{0em}
         \setlength{\leftskip}{1em}
         \medskip\noindent{\bfseries Solution to Problem {\theproblem}:}
     }
     {
%     \vskip-1.5em  \hphantom{a} \hfill \endex
      \medskip
       }
\newcommand{\res}[1]{Result~\ref{res:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\alg}[1]{Algorithm \ref{alg:#1}}
\renewcommand{\labelenumi}{(\alph{enumi})}

% Felisa's Shorthand notations
\def\points#1{{\bfseries [{#1} points]}\\%
}
\def\marks#1{{\bfseries [{#1} points]}}
\font\cmss=cmss10 scaled 1100
\font\goth=eufm10 at 11pt
\def\ds{\displaystyle}
\def\Field{{\hbox{\goth F}}}
\def\Filtration{{\mathbb F}}
\def\borel{{\cal B}}
\def\Real{{\mathbb R}}   % Real numbers
\def\Int{\mathbb N}  % Integer numbers
\def\Prob{{\mathbb P}}
\def\Esp{{\mathbb E}}
\def\prob{\hbox{\cmss P}}
\def\esp{\hbox{\cmss E}}   
\def\var{\hbox{\cmss Var}}   
\def\cov{\hbox{\cmss Cov}}   
\def\eqdist{{\,\buildrel \cal L\over = }\,}    
\def\todist{{\,\buildrel \cal L\over \Longrightarrow }\,}    
\def\eqdef{{\,\buildrel {\rm def} \over = }\,}
%
\def\normal{{\cal N}}  
\def\th{{\theta}}
\def\Th{{\Theta}}
\def\la{{\lambda}}
\def\ind#1{{\mathbf 1}_{\{#1\}}}
\def\given {\,|\,}
\def\order{{\cal O}}
\def\GG{{\cal G}}
\def\loss{{\cal L}}
\def\eff{{\cal E}}
\def\euro{ \hbox{\small{\hskip-.3em$\subset$ \hskip - 1.15 em  $-$}}}
\def\ep{{\epsilon}}
\newcommand\pfrac[2]{{\left(\frac{#1}{#2}\right)}}

\author{Morgan Wajda-Levie}
\date{October 20, 2022}



\begin{document}\sf

\lhead{ } \chead{ } \rhead{ } \lfoot{ } \cfoot{\sf -- page
\thepage~  of  \pageref{LastPage}  --  } \rfoot{ }                                                     %            <=====   PAGE NUMBERING

\title{\bfseries Computer Modeling  and  Simulation}

\maketitle

%\null \vskip - 5 cm

\begin{center}
\vskip - 1.5em { {\bfseries Problem Set 2}}
\end{center}

\begin{problem} \points{10}
Let  $\{X_n\}$ be a finite Markov Chain with state space $S$, and let $M=|S|<\infty$ be the number of states. Prove that if state $j$ is accessible from state $i$, then it is possible to attain state $j$ in at most $M$ time steps, starting at $X_0=i$. 
\end{problem}

\input{q1}

\begin{problem} \points{10}
Consider the  discrete queue model where at most a single customer arrives during a single service period, and the service time of a customer is a random variable $Z$ with the geometric probability distribution:

\[
\Prob(Z=k) = \alpha\, (1-\alpha)^{k-1}, \quad k=1,2,\ldots.
\]
Specify the transition probabilities of the Markov Chain whose state is the number of customers waiting for service or being served at the start of each period. Assume that the probability that a customer arrives in a period is $0<\beta<1$ and that no customer arrives with probability $1-\beta$.  
\end{problem}

\input{q2}

\begin{problem} \points{40}
This is a simulation of a typical station of public shared vehicles (such as the Citibike). The occupancy of the station at time $t$, denoted $X(t)$ is the number of vehicles parked. In this example we assume that all vehicles are ready to use (no need for repairs). Let $T=120$. During the interval of time $(0,T]$ the process evolves as follows, starting with $X(0)=10$. Bikes arrive at the station following a Poisson process with rate $\la = 6$. There are three types of clients that arrive requesting a vehicle according to independent Poisson process with respective rates $\mu_i, i=1,2,3$. Classes 1 and 2 are clients that pay annual memberships, and the revenue from these clients during $[0,T)$ is the aggregated (deterministic) amount $(K_1 \mu_1+ K_2\mu_2)T$, where $K_1 = .5, K_2=.1$. Class 3 describes clients that pay per ride an amount $K_3 = 1.25$. When annual members arrive at an empty station they are dissatisfied. This costs the system a penalty of $c_i, i=1,2$ where $c_1 = 1.00, c_2=0.25$. No penalty is paid for class 3 dissatisfactions. In your simulation, you will estimate the net profit during the interval of time $[0,T)$ using $\mu_1=3, \mu_2=1,\mu_3=4$.  
\begin{enumerate}
\item Describe the DES model and perform your simulations. Clarify how many replications of the interval $[0,T)$ you choose to use and how do you use these simulations in order to estimate the net profit. 
\item Now consider a retrospective model for the simulations instead. Use the following approach:
\begin{itemize}
\item[i)] Merge all Poisson processes. Name what theorems you use and what results from this merging. Describe the distribution of $M$, the total number of events in $[0,T)$. Given $M$, explain how you can calculate the times $(A_i, i=1,\ldots,M)$ of the $M$ events and name what theorems you use here. 
\item[ii)] Initialize your simulation generating $M$. Explain what method you use for this generation. 
\item[iii)] The rest of your code loops from $i=1$ to $M$. Explain how you calculate what type of event happens at every step in the code and name the theorems you use to justify this answer.
\end{itemize}
Clarify how many replications of the interval $[0,T)$ you choose to use and how do you use these simulations in order to estimate the net profit. 
\item Include a discussion comparing the two simulation approaches. 
\end{enumerate}
\end{problem}


\begin{problem}\points{20}
Consider the model of a M/GI/1 queue with Poisson arrivals of rate $\la$ (inter-arrival times are a sequence of iid exponential random variables with mean $1/\la$). Suppose that the service times $\{S_i\}$ are iid with general distribution. Call $Q(t)$ the {\em queue length} process at time $t$. It is well known that if $\rho\eqdef \la \Esp(S_1)<1$ then the stationary average queue length is:
\[
\theta =  \lim_{t\to\infty} \frac{1}{t} \int_0^t Q(s)\, ds = \frac{\la^2\,\Esp(S_1^2)}{2(1-\rho)}.
\]
\begin{enumerate}
\item Describe the discrete event model for the simulation: what are the variables in the physical state and what are the residual clocks? 
\item Write a code to estimate $\theta$ by simulating the queue length using the discrete event simulation model. Set $\la = 1$ and use the $\Gamma(3,0.25)$ distribution for the service times. Notice that this is an example of a  long term (stationary) average estimation. Explain how you determine the length of the simulation. 

\end{enumerate}
\end{problem}

\begin{problem} \points{20}
Consider again the queueing process of the previous problem. This time use the {\em sojourn time} process $\{X_n\}$, which for each customer $n$ provides the sum of wait plus service time: $X_n = W_n + S_n$. It is well known that fr the M/GI/1 model the sojourn time converges in distribution to a random  variable $X$ is, that is: $X_n \todist X$ as $n\to\infty$.  
Using Little's Law for the M/GI/1 queue of the previous problem, the stationary average {\em sojourn } time (waiting plus service) $X$ satisfies $\theta +\rho = \la X$, where $\th$ is the stationary queue length as in the previous problem. 
\begin{enumerate}
\item Describe the Petri Net model for simulation using Lindley's equation: what is the state of the system? What is the natural filtration that makes the model a Markov process (with continuous state space)?
\item Repeat (b) as in the previous problem, now using this simulation model instead (notice that the goal is to estimate $\th$ indirectly via the estimation of $X$).  
\end{enumerate}
\end{problem}

\end{document}
